---
title: "iTree Data Analysis_DNCC"
author: "COMPASS"
date: "12/14/2021"
output: word_document
---

# 1. Introduction
This document is a r-script for analysis of the tree data collected from Dhaka North City Corporation (DNCC) applying iTree Canopy platform. Species richness, tree density, basal area, tree cover, phtosociological attributes, diversity indices of the tree species, growing stock, biomass and carbon are the major variables estimated in this r-script using i-Tree data.There aare three strata and 11 land uses in DNCC. The key parameters were estimated for per strata and land uses.


# 2. Sampling process
The quadrat were 11.35 m radius circular plots.


# 3. Brief information regarding the databases used in this script
- plot = All plot basic data from i-Tree Eco Project
- tree = All tree data from i-Tree Eco Project
- gclu = Ground cover and land use data at plot level from the i-Tree Eco Project
- zone = Basic information regarding the three strata
- wd_sp_bd =  Wood density database from BFRI
- wd_sp = Wood density per species
- wd_gen = wood density per genus
- wd_fam = Wood density per family
- spp_info = family, local name and origin of the species recorded from DNCC


# 4. Data import
```{r}
# Remove memory
rm(list = ls())

# Set working directory
setwd("E:/Dropbox/iTree")


# Import survey data
zone <- read.csv("./data/stratum_info.csv", stringsAsFactors = F)
plot <- read.csv("./data/plot_v2_20220212.csv", stringsAsFactors = F)
tree <- read.csv("./data/tree_v3_20220211.csv", stringsAsFactors = F)
gclu <- read.csv("./data/gc_lu_20220211.csv", stringsAsFactors = F)

# Auxiliary data
spp_info <- read.csv("./data/added_tree_info.csv", stringsAsFactors = F)
wd_sp_bd <- read.csv("./data/wd_bd.csv", header=T, stringsAsFactors = F)
wd_fam <- read.csv("./data/wd_fam.csv", header=T, stringsAsFactors = FALSE)
wd_gen <- read.csv("./data/wd_gen.csv", header=T,  stringsAsFactors = FALSE)
wd_sp <- read.csv("./data/wd_sp.csv", header=T, stringsAsFactors = FALSE)


# Run the necessary libraries
library(data.table)
library(plotrix)
library(gtools)
require(reshape2)
library(ggplot2)
```


# 5. Data preparation
```{r}

# Data preparation
colnames(plot) <- c("plot_id", "stratum", "address", "lat", "long", "date",
                    "crew", "contact", "pa_ha", "ploto", "stake", "tree_p",
                    "shrub_p", "sampled_p", "notes", "complete")

colnames(tree) <- c("plot_id", "tree_id", "date", "status", "distance",
                    "direction", "species", "lu", "d1", "d1h", "d1m", "d2",
                    "d2h", "d2m", "d3", "d3h", "d3m", "d4", "d4h", "d4m", "d5",
                    "d5h", "d5m", "d6", "d6h", "d6m", "cd_p", "th", "cth",
                    "cbh", "cwns", "cwew", "cm_p", "i_p", "t_shrub_p", "cle",
                    "st", "public", "t_lat", "t_long", "t_comment")

# Consider the sampled plots only
plot <- plot[plot$complete=="TRUE",]

# Sampled area per plot
plot$ai <- (plot$sampled_p/100)*(pi*11.35^2)/10000


# Preparing tree cover data per plot
attach(plot)
plot$tree_p2 <- ifelse(tree_p=="100%", 100,
                       ifelse(tree_p=="95% - 99%", 97,
                              ifelse(tree_p=="90% - 95%", 92.5,
                                     ifelse(tree_p=="85% - 90%", 87.5,
                                            ifelse(tree_p=="80% - 85%", 82.5,
                                                   ifelse(tree_p=="75% - 80%", 77.5,
                                                          ifelse(tree_p=="70% - 75%", 72.5,
                                                                 ifelse(tree_p=="65% - 70%", 67.5,
                                                                        ifelse(tree_p=="60% - 65%", 62.5,
                                                                               ifelse(tree_p== "55% - 60%",57.5,
                                                                                      ifelse(tree_p=="50% - 55%", 52.5,
                                                                                             ifelse(tree_p=="45% - 50%", 47.5,
                                                                                                    ifelse(tree_p=="40% - 45%", 42.5,
ifelse(tree_p=="35% - 40%", 37.5,
       ifelse(tree_p=="30% - 35%", 32.5,
              ifelse(tree_p=="25% - 30%", 27.5,
                     ifelse(tree_p=="20% - 25%", 22.5,
                            ifelse(tree_p=="15% - 20%", 17.5,
                                   ifelse(tree_p=="10% - 15%", 12.5,
                                          ifelse(tree_p=="5% - 10%",
                                                 7.5, ifelse(tree_p=="1% - 5%", 3,
                                                             ifelse(tree_p=="0%", 0, "Not Entered"))))))))))))))))))))))


# Add ground cover and land use with plot
#plot2 <- merge(plot, gclu, by="plot_id", all.x=T)

# Add additional species information with the tree data base
tree <- merge(tree, spp_info, by="species", all.x=T)
tree$count <- 1

# Add stratum name with tree
tree <- merge(tree, plot[,c("plot_id", "stratum", "ai")], by="plot_id", all.x=T)
tree <- tree[!is.na(tree$stratum),]

```



# 6. Species richness
```{r}
# Species composition
sp_comp <- unique(tree[,c("species", "family", "exotic_nat")])
write.csv(sp_comp, file="./analysis_outputs/1. Species composition.csv", row.names = F)
```


## 6.1 Species richness by stratum
```{r}
# Species number per stratum
t_z <- unique(tree[,c("stratum", "species", "local_name", "family", "exotic_nat", "count")]) # t_z = Number of tree species per stratum
t_z$exotic <- ifelse(t_z$exotic_nat=="Exotic", 1, 0)
t_z$native <- ifelse(t_z$exotic_nat=="Native", 1, 0)
t_z <- data.table(t_z)
t_z2 <- t_z[,list(spp_no = length(unique(species)),
               fam_no = length(unique(family)),
               native = sum(native),
               exotic = sum(exotic)),
            by="stratum"]

# Export output
write.csv(t_z2, file="./analysis_outputs/3.1.1. Number of species per stratum.csv", row.names = F)

# Number of species in all zone
length(unique(tree$species))
# Number of families in all zone
length(unique(tree$family))
# Number of exotic species
length(unique(tree[,c("species", "exotic_nat")][tree$exotic_nat=="Exotic",])$exotic_nat)
# Number of native species
length(unique(tree[,c("species", "exotic_nat")][tree$exotic_nat=="Native",])$exotic_nat)


```


## 6.2 Species richness by land use
```{r}
# Species number per land use
t_lu <- unique(tree[,c("lu", "species", "local_name", "family", "exotic_nat", "count")]) 
t_lu$exotic <- ifelse(t_lu$exotic_nat=="Exotic", 1, 0)
t_lu$native <- ifelse(t_lu$exotic_nat=="Native", 1, 0)
t_lu <- data.table(t_lu)
t_lu2 <- t_lu[,list(spp_no = length(unique(species)),
               fam_no = length(unique(family)),
               native = sum(native),
               exotic = sum(exotic)),
            by="lu"]

# Export output
write.csv(t_lu2, file="./analysis_outputs/3.1.2. Number of species per land use.csv", row.names = F)


```


# 7. Phytosociological attributes
## List of variable names prepared and usind in the data frame for anlysis
- a = number of individuals of the ith species occurred in the area of interest;
- b = total number of quadrats studied;
- c = total number of quadrats in which the species occurred;
- M = Total number of individuals of all the species in the area of interest,
- F = frequency of ith species, S = number of species.
- ai_s = Sampled area per stratum

## 7.1 Phytosociological attributes per stratum
```{r}
# Number of individuals (a) per species and stratum
PA_s <- aggregate(data=tree, count~stratum+species, sum)
names(PA_s)[3] <- "a"
# Add the sampled area and number of quadrats per stratum
PA_s <- merge(PA_s, zone[,c("stratum", "sampled_plot_no", "sampled_area_ha")], by= "stratum", all.x = T)
names(PA_s)[names(PA_s)=="sampled_plot_no"] <- "b"
names(PA_s)[5] <- "ai_s"

# Number of plots occurred per species and stratum
npo <- unique(tree[,c("plot_id", "stratum", "species", "count")])
np_ss <- aggregate(data=npo, count~species+stratum, sum) #np_ss = no. of plots occurred by species and stratum
names(np_ss)[3] <- "c"
PA_s <- merge(PA_s, np_ss, by=c("stratum", "species"), all.x=T)

# Total number of individuals (M) of all species per stratum
ni_s <- aggregate(data=PA_s, a~stratum, sum) #Number of individuals of all species per site
names(ni_s)[2] <- "M"
PA_s <- merge(PA_s, ni_s, by="stratum", all.x=T)

# Add the number of species per stratum from "t_z2" data frame
PA_s <- merge(PA_s, t_z2[,c("stratum", "spp_no")], by="stratum", all.x = T)
names(PA_s)[8] <- "S"

# Stem density per ha
PA_s$td_ha <- PA_s$a/PA_s$ai_s

# Density
PA_s$D <- PA_s$a/PA_s$b
PA_s$RD <- (PA_s$a/PA_s$M)*100

# Frequency
PA_s$F <- PA_s$c/PA_s$b
PA_s$RF <- ifelse(PA_s$stratum=="Tree", (PA_s$F/sum(PA_s[PA_s$stratum=="Tree",]$F))*100,
                    ifelse(PA_s$stratum=="Scatter", (PA_s$F/sum(PA_s[PA_s$stratum=="Scatter",]$F))*100,
                           ifelse(PA_s$stratum=="Others", (PA_s$F/sum(PA_s[PA_s$stratum=="Others",]$F))*100, NA)))

# Abundabce
PA_s$A <- PA_s$a/PA_s$c
PA_s$RA <- ifelse(PA_s$stratum=="Tree", (PA_s$A/sum(PA_s[PA_s$stratum=="Tree",]$A))*100,
                    ifelse(PA_s$stratum=="Scatter", (PA_s$A/sum(PA_s[PA_s$stratum=="Scatter",]$A))*100,
                           ifelse(PA_s$stratum=="Others", (PA_s$A/sum(PA_s[PA_s$stratum=="Others",]$A))*100, NA)))

# Importance Vlaue index (IVI)
PA_s$IVI <- PA_s$RD+PA_s$RF+PA_s$RA

# Export the output
write.csv(PA_s, file="./analysis_outputs/Phytosociological attributes per statum.csv", row.names = F)
```


## 7.2 Phytosociological attributes for DNCC
```{r}
# Stem density per ha per species
PA_s  <- merge(PA_s, zone[,c("stratum", "Wh")], by="stratum", all.x=T)
PA_s$wtd_ha <- PA_s$td_ha*PA_s$Wh #wtd_ha = weighted tree density per ha
td_sp <- aggregate(data=PA_s, wtd_ha~species, sum)

# Number of individuals (a) per species
PA <- aggregate(data=tree, count~species, sum)
names(PA)[2] <- "a"
# Add the total number of quadrat studied
PA$b <- sum(zone$sampled_plot_no)

# Number of plots occurred per species
npo_sp <- unique(tree[,c("plot_id", "species", "count")])
np_sp <- aggregate(data=npo_sp, count~species, sum) #np_sp = no. of plots occurred by species
names(np_sp)[2] <- "c"
PA <- merge(PA, np_sp, by="species", all.x=T)

# Total number of individuals (M) of all species
PA$M <- sum(PA$a)

# Add stem density per ha
PA <- merge(PA, td_sp, by="species", all.x=T)

# Density
PA$D <- PA$a/PA$b
PA$RD <- (PA$a/PA$M)*100

# Frequency
PA$F <- PA$c/PA$b
PA$RF <- (PA$F/sum(PA$F))*100

# Abundabce
PA$A <- PA$a/PA$c
PA$RA <- PA$A/sum(PA$A)*100

# Importance Vlaue index (IVI)
PA$IVI <- PA$RD+PA$RF+PA$RA

# Export the output
write.csv(PA, file="./analysis_outputs/Phytosociological attributes for DNCC.csv", row.names = F)
            
```


# 8. Diversity indicers of the tree species
Four diversity indices were calculated.

## 8.1 Diversity indicers of the tree species per stratum
```{r}
DI_s <- PA_s[,c("species", "stratum", "a", "M")]
DI_s$Pi <- DI_s$a/DI_s$M
DI_s$Pi2 <- DI_s$Pi^2
DI_s$Pi_LnPi <- DI_s$Pi*log(DI_s$Pi)
DI_s$aiai_1 <- DI_s$a*(DI_s$a-1) #aiai_1 = (a(a-1)) of the equation

# Prepare the diversity indices per stratum table
DI_sF <- data.table(DI_s)
DI_sF <- DI_sF[,list(H = sum(-Pi_LnPi),
                     Sd1 = sum(aiai_1)), by="stratum"]

# Add number of species and total number of individuals per stratum
DI_sF <- merge(DI_sF, unique(PA_s[,c("stratum", "M", "S")]), by="stratum", all.x=T)
DI_sF$R <- (DI_sF$S-1)/log(DI_sF$M) # R = Richness index
DI_sF$Sd <- 1-(DI_sF$Sd1)/(DI_sF$M*(DI_sF$M-1)) # Sd = Simpson's Dominance Index
DI_sF$E <- DI_sF$H/log(DI_sF$S) # E = Pilou's evenness index

# Export the outptus
write.csv(DI_sF, file="./analysis_outputs/Diversity indices per stratum.csv", row.names = F)

```


## 8.2 Diversity indicers for whole DNCC
```{r}
DI <- aggregate(data=tree, count~species, sum)
names(DI)[2] <- "a"
DI$M <- sum(DI$a)
DI$Pi <- DI$a/DI$M
DI$Pi2 <- DI$Pi^2
DI$Pi_LnPi <- DI$Pi*log(DI$Pi)
DI$aiai_1 <- DI$a*(DI$a-1)

# Prepare the diversity indices per stratum table
DIF <- data.frame(H = sum(-DI$Pi_LnPi),
                  R = (length(unique(tree$species))-1)/log(sum(DI$a)),
                  Sd = 1-(sum(DI$aiai_1)/(sum(DI$a)*(sum(DI$a)-1))),
                  E = sum(-DI$Pi_LnPi)/log(length(unique(tree$species))))



# Export the outptus
write.csv(DIF, file="./analysis_outputs/Diversity indices for whole DNCC.csv", row.names = F)

```


# 9. Estimating the key variables at tree level
## 9.1 Basal Area
```{r}
tree$d1_m <- tree$d1/100
tree$d2_m <- tree$d2/100
tree$d3_m <- tree$d3/100
tree$d4_m <- tree$d4/100
tree$d5_m <- tree$d5/100
tree$d6_m <- tree$d6/100

tree$ba1 <- (pi*(tree$d1_m^2))/4 #ba1 = Basal area based on diameter 1 in m2
tree$ba2 <- (pi*(tree$d2_m^2))/4
tree$ba3 <- (pi*(tree$d3_m^2))/4
tree$ba4 <- (pi*(tree$d4_m^2))/4
tree$ba5 <- (pi*(tree$d5_m^2))/4
tree$ba6 <- (pi*(tree$d6_m^2))/4

which(colnames(tree)=="ba1") #To know the column number
match("ba6", names(tree)) #To know the column number

tree[,57:62][is.na(tree[,57:62])] <- 0 # Convert NAs into 0
tree$ba <- tree$ba1+tree$ba2+tree$ba3+tree$ba4+tree$ba5+tree$ba6


```



## 9.2 Growing Stock
Growing stock is estimated using the quarter girth equation with an average form factor (0.693) applicable for Bangladesh. The form factor was adopted from BFI Cycle 1 data analysis method.  
```{r}
tree$gs1 <- ifelse(tree$d1>=10, (pi*(tree$d1_m^2)*tree$th*0.693)/4, 0) #gs1 = Growing stock in m3 based on DBH 1
tree$gs2 <- ifelse(tree$d2>=10, (pi*(tree$d2_m^2)*tree$th*0.693)/4, 0) #gs2 = Growing stock in m3 based on DBH 2
tree$gs3 <- ifelse(tree$d3>=10, (pi*(tree$d3_m^2)*tree$th*0.693)/4, 0) #gs3 = Growing stock in m3 based on DBH 3
tree$gs4 <- ifelse(tree$d4>=10, (pi*(tree$d4_m^2)*tree$th*0.693)/4, 0) #gs4 = Growing stock in m3 based on DBH 4
tree$gs5 <- ifelse(tree$d5>=10, (pi*(tree$d5_m^2)*tree$th*0.693)/4, 0) #gs5 = Growing stock in m3 based on DBH 5
tree$gs6 <- ifelse(tree$d6>=10, (pi*(tree$d6_m^2)*tree$th*0.693)/4, 0) #gs6 = Growing stock in m3 based on DBH 6

# gs1 column number
which(colnames(tree)== "gs1")
match("gs6", names(tree)) #To know the column number

# Convert NAs into zero
tree[,64:69][is.na(tree[,64:69])] <- 0
tree$gs <- tree$gs1+tree$gs2+tree$gs3+tree$gs4+tree$gs5+tree$gs6

```


## 9.3 Biomass
### 9.3.1 Above Ground Biomass
Above Ground Biomss (AGB) is estimated using the pantropical allomet Chaves et all 2014.
```{r}
# Adding wood density data
treeB <- merge(tree, wd_sp_bd[,c("tree_spp_scientific_name", "wd_bd")], by.x="genus_species",
               by.y="tree_spp_scientific_name", all.x = T)
treeB <- merge(treeB, wd_sp[,c("Binomial","wd_sp")], by.x = "genus_species", by.y="Binomial",
               all.x =TRUE)
treeB <- merge(treeB, wd_gen[,c("Genus","wd_gen")], by.x = "genus_species", by.y="Genus",
               all.x =TRUE)
treeB <- merge(treeB, wd_fam[,c("Family","wd_fam")],by.x = "family", by.y="Family", all.x=TRUE)

# Prepare wd column with wd from different columns
treeB$wd <- ifelse(!is.na(treeB$wd_bd), treeB$wd_bd,
                   ifelse(!is.na(treeB$wd_sp), treeB$wd_sp,
                          ifelse(!is.na(treeB$wd_gen), treeB$wd_gen,
                                 ifelse(!is.na(treeB$wd_fam), treeB$wd_fam,
                                        0.6133782)))) # wd= wood density in g/cm3
# Checking NA in wd column
treeB_NAWD <-treeB[is.na(treeB$wd),]

#Applying Chaves et al. (2014) best-fitted pantropical equation for above ground biomass (AGB) estimation
treeB$agb1 <- 0.0673*(treeB$wd*(treeB$d1^2)*treeB$th)^0.976 #abg1 = AGB in kg for DBH 1
treeB$agb2 <- 0.0673*(treeB$wd*(treeB$d2^2)*treeB$th)^0.976 #abg2 = AGB in kg for DBH 2
treeB$agb3 <- 0.0673*(treeB$wd*(treeB$d3^2)*treeB$th)^0.976 #abg3 = AGB in kg for DBH 3
treeB$agb4 <- 0.0673*(treeB$wd*(treeB$d4^2)*treeB$th)^0.976 #abg4 = AGB in kg for DBH 4
treeB$agb5 <- 0.0673*(treeB$wd*(treeB$d5^2)*treeB$th)^0.976 #abg5 = AGB in kg for DBH 5
treeB$agb6 <- 0.0673*(treeB$wd*(treeB$d6^2)*treeB$th)^0.976 #abg6 = AGB in kg for DBH 6

# Check agb1 and agb6 colunm number
which(colnames(treeB)=="agb1")
match("agb6", names(treeB)) #To know the column number

# Convert NA in agb1 to agb6 into zero
treeB[,76:81][is.na(treeB[,76:81])] <- 0

# agb in kg per tree
treeB$agb <- treeB$agb1+treeB$agb2+treeB$agb3+treeB$agb4+treeB$agb5+treeB$agb6
# agb in ton
treeB$agb_t <- treeB$agb/1000

```


### 9.3.2 Below ground biomass
Below Ground Biomass (BGB) is calculated using Pearson et al. (2007) equation for BGB. Accoriding to Pearson et al. (2007), BGB = exp(-1.0587 + 0.8836 x ln AGB).

```{r}
treeB$bgb_t <- exp(-1.0587 + 0.8836*log(treeB$agb_t))
```

### 9.3.3 Total tree biomass
```{r}
treeB$ttb_t <- treeB$agb_t+treeB$bgb_t
```



## 9.4. Carbon
Carbon in AGB and BGB is estimated as 50% of the AGB and BGB respectively.
```{r}
# A separate file for carbon
treeC <- treeB
# Carbon in AGB
treeC$cagb_t <- treeC$agb_t*0.5
# Carbon in BGB
treeC$cbgb_t <- treeC$bgb_t*0.5
# Total carbon per tree
treeC$carbon_t <- treeC$cagb_t+treeC$cbgb_t

```


# 10 Estimating means at strata and DNCC level

Under this section the key variables- stem density, basal area, tree cover, growing stock, AGB, BGB, TTB, CAGB, CBGB, and TTC were estimated per stratum and whole DNCC.

```{r}
# All key variable at plot level
kv_p <- data.table(treeC)
kv_p <- kv_p[,list(nt_p = sum(count),
                   ba_p = sum(ba),
                   gs_p = sum(gs),
                   agb_p_t = sum(agb_t),
                   bgb_p_t = sum(bgb_t),
                   ttb_p_t = sum(ttb_t),
                   cagb_p_t = sum(cagb_t),
                   cbgb_p_t = sum(cbgb_t),
                   ttc_p_t = sum(carbon_t)),
             by = "plot_id"]  #kv_p = key variabels per plot

# add sampled area and tc of all measured plots
kv_p2 <- merge(kv_p, plot[,c("plot_id", "stratum", "tree_p2", "ai")], by="plot_id", all.y=T)

# Convert NAs into zero
kv_p2[,2:10][is.na(kv_p2[,2:10])] <- 0

# Calculate the variables per hectare basis
kv_p2$nt_p_ha <- kv_p2$nt_p/kv_p2$ai
kv_p2$ba_p_ha <- kv_p2$ba_p/kv_p2$ai
kv_p2$gs_p_ha <- kv_p2$gs_p/kv_p2$ai
kv_p2$agb_p_tha <- kv_p2$agb_p_t/kv_p2$ai
kv_p2$bgb_p_tha <- kv_p2$bgb_p_t/kv_p2$ai
kv_p2$ttb_p_tha <- kv_p2$ttb_p_t/kv_p2$ai
kv_p2$cagb_p_tha <- kv_p2$cagb_p_t/kv_p2$ai
kv_p2$cbgb_p_tha <- kv_p2$cbgb_p_t/kv_p2$ai
kv_p2$ttc_p_tha <- kv_p2$ttc_p_t/kv_p2$ai

# Convert tree cover as numeric
kv_p2$tree_p2 <- as.numeric(kv_p2$tree_p2)

# Key variable at stratum level
kv_s <- data.table(kv_p2)
kv_s <- kv_s[,list(nt_ha = mean(nt_p_ha),
                   nt_se = std.error(nt_p_ha),
                   ba_ha = mean(ba_p_ha),
                   ba_se = std.error(ba_p_ha),
                   tc = mean(tree_p2),
                   tc_se = std.error(tree_p2),
                   gs_ha = mean(gs_p_ha),
                   gs_se = std.error(gs_p_ha),
                   agb_ha = mean(agb_p_tha),
                   agb_se = std.error(agb_p_tha),
                   bgb_ha = mean(bgb_p_tha),
                   bgb_se = std.error(bgb_p_tha),
                   ttb_ha = mean(ttb_p_tha),
                   ttb_se = std.error(ttb_p_tha),
                   cagb_ha = mean(cagb_p_tha),
                   cagb_se = std.error(cagb_p_tha),
                   cbgb_ha = mean(cbgb_p_tha),
                   cbgb_se = std.error(cbgb_p_tha),
                   ttc_ha = mean(ttc_p_tha),
                   ttc_se = std.error(ttc_p_tha)),
             by="stratum"]

# Add stratum weight
kv_s2 <- merge(kv_s, zone[,c("stratum", "Wh")], by = "stratum", all.x=T)

# multiply key variables with stratum weights
kv_s2$nt_ha_w <- kv_s2$nt_ha*kv_s2$Wh
kv_s2$ba_ha_w <- kv_s2$ba_ha*kv_s2$Wh
kv_s2$tc_w <- kv_s2$tc*kv_s2$Wh
kv_s2$gs_ha_w <- kv_s2$gs_ha*kv_s2$Wh
kv_s2$agb_ha_w <- kv_s2$agb_ha*kv_s2$Wh
kv_s2$bgb_ha_w <- kv_s2$bgb_ha*kv_s2$Wh
kv_s2$ttb_ha_w <- kv_s2$ttb_ha*kv_s2$Wh
kv_s2$cagb_ha_w <- kv_s2$cagb_ha*kv_s2$Wh
kv_s2$cbgb_ha_w <- kv_s2$cbgb_ha*kv_s2$Wh
kv_s2$ttc_ha_w <- kv_s2$ttc_ha*kv_s2$Wh

# Key avraibels at DNCC levels
kv_DNCC <- data.frame(nt_ha = sum(kv_s2$nt_ha_w),
                      nt_se = std.error(kv_p2$nt_p_ha),
                      ba_ha = sum(kv_s2$ba_ha_w),
                      ba_se = std.error(kv_p2$ba_p_ha),
                      tc = sum(kv_s2$tc_w),
                      tc_se = std.error(kv_p2$tree_p2),
                      gs_ha = sum(kv_s2$gs_ha_w),
                      gs_se = std.error(kv_p2$gs_p_ha),
                      agb_ha = sum(kv_s2$agb_ha_w),
                      agb_se = std.error(kv_p2$agb_p_tha),
                      bgb_ha = sum(kv_s2$bgb_ha_w),
                      bgb_se = std.error(kv_p2$bgb_p_tha),
                      ttb_ha = sum(kv_s2$ttb_ha_w),
                      ttb_se = std.error(kv_p2$ttb_p_tha),
                      cagb_ha = sum(kv_s2$cagb_ha_w),
                      cagb_se = std.error(kv_p2$cagb_p_tha),
                      cbgb_ha = sum(kv_s2$cbgb_ha_w),
                      cbgb_se = std.error(kv_p2$cbgb_p_tha),
                      ttc_ha = sum(kv_s2$ttc_ha_w),
                      ttc_se = std.error(kv_p2$ttc_p_tha))

kv_DNCC$stratum <- "DNCC"

# Combine stratum and DNCC level estimates
kv_F <- smartbind(kv_s2, kv_DNCC)
# kv_F = rbind.fill(kv_s2, kv_DNCC) #library("plyr")

# Export the outputs
write.csv(kv_F[,c("stratum", "nt_ha", "nt_se", "ba_ha", "ba_se", "tc", "tc_se", "gs_ha", "gs_se", "agb_ha", "agb_se", "bgb_ha", "bgb_se", "ttb_ha", "ttb_se", "cagb_ha", "cagb_se", "cbgb_ha", "cbgb_se", "ttc_ha", "ttc_se")], file = "./analysis_outputs/key variables per stratum and DNCC.csv", row.names = F)

```


# 11. Estimating means of the key variables per land use and DNCC level

Under this section the key variables- stem density, basal area, tree cover, growing stock, AGB, BGB, TTB, CAGB, CBGB, and TTC were estimated per land uses (identified by the field teams) and whole DNCC.

```{r}
# prepare the land use table
# Plot area in ha
pa <- (pi*11.35^2)/10000 # pa = plot area in ha

# LU area (ha) per plot
gclu$LU_vacant_ha <- (gclu$LU_vacant/100)*pa
gclu$LU_residential_ha <- (gclu$LU_residential/100)*pa
gclu$LU_forest_ha <- (gclu$LU_forest/100)*pa
gclu$LU_water_wetland_ha <- (gclu$LU_water_wetland/100)*pa
gclu$LU_park_ha <- (gclu$LU_park/100)*pa
gclu$LU_cemetery_ha <- (gclu$LU_cemetery/100)*pa
gclu$LU_transportation_ha <- (gclu$LU_transportation/100)*pa
gclu$LU_agriculture_ha <- (gclu$LU_agriculture/100)*pa
gclu$LU_institutional_ha <- (gclu$LU_institutional/100)*pa
gclu$LU_commercial_industrial_ha <- (gclu$LU_commercial_industrial/100)*pa
gclu$LU_other_ha <- (gclu$LU_other/100)*pa



# Add complete status
gclu <- merge(gclu, plot[,c("plot_id", "complete")], by="plot_id", all.x=T)

# Select only completely measured plots
gclu <- gclu[!is.na(gclu$complete),]


# Total sampled area (ha) per land use
ai_lu <- data.frame(Vacant = sum(gclu$LU_vacant_ha),
                    Residential = sum(gclu$LU_residential_ha),
                    Forest = sum(gclu$LU_forest_ha),
                    Water_Wetland = sum(gclu$LU_water_wetland_ha),
                    Park = sum(gclu$LU_park_ha),
                    Cemetery = sum(gclu$LU_cemetery_ha),
                    Transportation = sum(gclu$LU_transportation_ha),
                    Agriculture = sum(gclu$LU_agriculture_ha),
                    Institutional = sum(gclu$LU_institutional_ha),
                    Commercial_Industrial = sum(gclu$LU_commercial_industrial_ha),
                    Other = sum(gclu$LU_other_ha))
# Melt the area per lu
# Create ID 
ai_lu$lu <- "ai_lu_ha"
ai_lu_molted <- melt(ai_lu, "lu")
# rename columns
names(ai_lu_molted)[2:3] <- c("lu", "ai_lu_ha")
ai_luF <- ai_lu_molted[,2:3]



# Key variables per lu
kv_lu <- data.table(treeC)
kv_lu <- kv_lu[,list(nt_lu = sum(count),
                   ba_lu = sum(ba),
                   gs_lu = sum(gs),
                   agb_lu_t = sum(agb_t),
                   bgb_lu_t = sum(bgb_t),
                   ttb_lu_t = sum(ttb_t),
                   cagb_lu_t = sum(cagb_t),
                   cbgb_lu_t = sum(cbgb_t),
                   ttc_lu_t = sum(carbon_t)),
             by = "lu"]  #kv_lu = key variabels per land use
# Rename "Commercial/Industrial" land use
kv_lu$lu <- ifelse(kv_lu$lu=="Commercial/Industrial", "Commercial_Industrial", kv_lu$lu)

# Add sampled area per lu
kv_lu2 <- merge(kv_lu, ai_luF, by="lu", all.x=T)

# Estimate the variables per hectare basis
kv_lu2$nt_ha <- kv_lu2$nt_lu/kv_lu2$ai_lu_ha
kv_lu2$ba_ha <- kv_lu2$ba_lu/kv_lu2$ai_lu_ha
kv_lu2$gs_ha <- kv_lu2$gs_lu/kv_lu2$ai_lu_ha
kv_lu2$agb_tha <- kv_lu2$agb_lu_t/kv_lu2$ai_lu_ha
kv_lu2$bgb_tha <- kv_lu2$bgb_lu_t/kv_lu2$ai_lu_ha
kv_lu2$ttb_tha <- kv_lu2$ttb_lu_t/kv_lu2$ai_lu_ha
kv_lu2$cagb_tha <- kv_lu2$cagb_lu_t/kv_lu2$ai_lu_ha
kv_lu2$cbgb_tha <- kv_lu2$cbgb_lu_t/kv_lu2$ai_lu_ha
kv_lu2$ttc_tha <- kv_lu2$ttc_lu_t/kv_lu2$ai_lu_ha

# export the outputs
write.csv(kv_lu2[,c(1,12:20)], file="./analysis_outputs/key variables per land use.csv", row.names = F)


# Prepare graphs

## 3.2.2 Tree density per land use
p322_td <- ggplot(data = kv_lu2, aes(x = lu, y = nt_ha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Tree~density~(stem~ha^-1)))+
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.2.2. Tree density per land use.tiff", p322_td, units="in", width=3.2,
         height=4, dpi=300, compression = 'lzw')

## 3.2.2 Tree basal area per land use
p322_ba <- ggplot(data = kv_lu2, aes(x = lu, y = ba_ha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Basal~area~m^2~ha^-1))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.2.2. Basal area per land use.tiff", p322_ba, units="in", width=3.2,
         height=4, dpi=300, compression = 'lzw')

## 3.5.2 Growing stock per land use
p352 <- ggplot(data = kv_lu2, aes(x = lu, y = gs_ha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Growing~stock~m^3~ha^-1))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.5.2. Growing stock per land use.tiff", p352, units="in", width=6.4,
         height=5, dpi=300, compression = 'lzw')



## 3.6.2 Above Ground Biomass per land use
p362_agb <- ggplot(data = kv_lu2, aes(x = lu, y = agb_tha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Above~ground~biomass~(t~ha^-1)))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.6.2. AGB per land use.tiff", p362_agb, units="in", width=3.2,
         height=4, dpi=300, compression = 'lzw')


## 3.6.2 Below Ground Biomass per land use
p362_bgb <- ggplot(data = kv_lu2, aes(x = lu, y = bgb_tha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Below~ground~biomass~(t~ha^-1)))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.6.2. BGB per land use.tiff", p362_bgb, units="in", width=3.2,
         height=4, dpi=300, compression = 'lzw')


## 3.7.2 Carbon in Above Ground Biomass per land use
p372_cagb <- ggplot(data = kv_lu2, aes(x = lu, y = cagb_tha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Carbon~"in"~AGB~(t~ha^-1)))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.7.2. CAGB per land use.tiff", p372_cagb, units="in", width=3.2,
         height=3.7, dpi=300, compression = 'lzw')


## 3.7.2 Carbon in Below Ground Biomass per land use
p372_cbgb <- ggplot(data = kv_lu2, aes(x = lu, y = cbgb_tha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Carbon~"in"~BGB~(t~ha^-1)))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.7.2. CBGB per land use.tiff", p372_cbgb, units="in", width=3.2,
         height=3.7, dpi=300, compression = 'lzw')


## 3.7.2 Total Tree Carbon per land use
p372_ttc <- ggplot(data = kv_lu2, aes(x = lu, y = ttc_tha))+
  geom_bar(stat = "identity")+
  xlab("Land uses")+ylab(expression(Total~tree~carbon~(t~ha^-1)))+ #expression for super script
  theme(axis.line = element_line(colour = "black", size = 0.5, linetype = "solid"))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggsave("./analysis_outputs/3.7.2. Total tree carbon per land use.tiff", p372_ttc, units="in", width=6.4, height=4.3, dpi=300, compression = 'lzw')

```


