This R code contains scripts for the analysis presented in the follwing paper.

Majumder, S.C., Uddin, M.S., Hossain, M.A., Iqbal, M.Z., Sarkar, P.R., Johnson, K., Aziz, M.T., Rahaman, M.T., Zzaman, M.B., Hossain, M.K., Islam, K.K., Rony, A., Shakil, S.H., Haque, M.A., Hossain, M.Z., Henning, J., Paqueo, L.V.C., Green, J., 2026. Urban forests baseline and ecosystem benefits of a tropical metropolis: case of Dhaka, Bangladesh. Sci. Rep. https://doi.org/10.1038/s41598-026-46985-x


# Set WD
```{r}
setwd("C:/OneDrive/iTree")
```


# Import data
```{r}
zone <- read.csv("./data/stratum_info.csv", stringsAsFactors = F)
plot <- read.csv("./data/Plot17112021.csv", stringsAsFactors = F)
tree <- read.csv("./data/tree17112021_v2.csv", stringsAsFactors = F)
```

# Run the necessary libraries
```{r}
library(data.table)
library(plotrix)
```



# Data preparation
```{r}
colnames(plot) <- c("plot_id", "stratum", "address", "lat", "long", "date", "crew", "contact", "pa_ha", "ploto", "stake", "tree_p", "shrub_p", "sampled_p", "notes", "complete")

colnames(tree) <- c("plot_id", "tree_id", "date", "status", "distance", "direction", "species", "lu", "d1", "d1h", "d1m", "d2", "d2h", "d2m", "d3", "d3h", "d3m", "d4", "d4h", "d4m", "d5", "d5h", "d5m", "d6", "d6h", "d6m", "cd_p", "th", "cth", "cbh", "cwns", "cwew", "cm_p", "i_p", "t_shrub_p", "cle", "st", "public", "t_lat", "t_long", "t_comment")

# Consider the sampled plots only
plot <- plot[plot$complete=="True",]

# Sampled area per plot
plot$ai <- (plot$sampled_p/100)*(pi*11.35^2)/10000
```

